On the Design, Implementation and Use of Laziness in R

### 2019-08-15 report/paper-extra.Rmd
Update reports and add new analysis

@@

# Latex Macro Generation
## name prefix is one of:
## - Perc - adds a % at the end
## - Count - adds labels like K, M, B for Kilo, Million and Billion
## - Size - adds lables like KB, MB, GB for memory sizes
generate_macro <- function(macro_name, value, subst = TRUE) {
    if(subst) {
        name <- as.character(substitute(macro_name))
    } else {
        name <- macro_name
    }
    if (str_sub(name, 1, 4) == "Perc")      formatted_value <- format_as_percentage(value)
    else if(str_sub(name, 1, 5) == "Count") formatted_value <- format_as_absolute(value)
    else if(str_sub(name, 1, 4) == "Size")  formatted_value <- format_as_size(value)
    else stop(str_c("Unidentified name", name, sep = " "))
    formatted_value <- str_replace(formatted_value, "\\%", "\\\\%")
    definition <- str_glue("\\newcommand {{\\{name}}} {{{formatted_value}\\xspace}} %% {value}")
    write_lines(definition, params$latex_macro_filepath, append = TRUE)
}

@@

promise_argument_forced_by_promise_argument <- read_table("promise_argument_forced_by_promise_argument")
promise_argument_returning_non_locally <- read_table("promise_argument_returning_non_locally")
```

@@

@@

```{r, eval = TRUE, echo = FALSE}
datatable(promise_lifecycle_summary)

promise_lifecycle_list %>%
filter(argument, escaped) %>%
datatable()

promise_lifecycle_list %>%
filter(!argument, escaped) %>%
datatable()

promise_lifecycle_list %>%
filter(argument, !escaped) %>%
datatable()

promise_lifecycle_list %>%
filter(!argument, !escaped) %>%
datatable()
```

@@

datatable(context_sensitive_lookup_symbols)
```

@@

@@

```{r, eval = TRUE, echo = FALSE}
print("TODO")
```

@@

```{r, eval = TRUE, echo = FALSE}
datatable(argument_promise_count_by_call_depth)

@@

save_local_graph(argument_promise_call_depth_graph, "argument_promise_call_depth_graph.pdf", width=12, height=2.5)
```

@@

```{r, eval = TRUE, echo = FALSE}
datatable(promise_argument_forced_by_promise_argument)
```

@@

```{r, eval = TRUE, echo = FALSE}
datatable(argument_promise_count_by_forcing_scope)
```

>>>>

format_as_code <- function(x) paste0("\\code{", x,  "}")

@@

# Latex Macro Generation
## name prefix is one of:
## - Perc - adds a % at the end
## - Count - adds labels like K, M, B for Kilo, Million and Billion
## - Size - adds lables like KB, MB, GB for memory sizes
## - Code - wraps the value in \code{}
## - Text - content is as is
generate_macro <- function(macro_name, value, subst = FALSE) {
    if(subst) {
        name <- as.character(substitute(macro_name))
    } else {
        name <- macro_name
    }
    if (str_sub(name, 1, 4) == "Perc")      formatted_value <- format_as_percentage(value)
    else if(str_sub(name, 1, 5) == "Count") formatted_value <- format_as_absolute(value)
    else if(str_sub(name, 1, 4) == "Size")  formatted_value <- format_as_size(value)
    else if(str_sub(name, 1, 4) == "Code")  formatted_value <- format_as_code(value)
    else if(str_sub(name, 1, 4) == "Text")  formatted_value <- format_as_code(value)
    else stop(str_c("Unidentified name", name, sep = " "))
    formatted_value <- str_replace(formatted_value, "\\%", "\\\\%")
    definition <- str_glue("\\newcommand {{\\{name}}} {{{formatted_value}\\xspace}} %% {value}")
    write_lines(definition, params$latex_macro_filepath, append = TRUE)
}

@@

promise_argument_forced_by_promise_argument <- read_table("promise_argument_forced_by_promise_argument")
promise_argument_returning_non_locally <- read_table("promise_argument_returning_non_locally")

created_side_effects <- read_table("created_side_effects")
observed_side_effects <- read_table("observed_side_effects")
```

@@

PercCallerEnvSubstituteCalls <-
    substitute_class_summary %>%
    filter(substitute_class == "SameScope") %>%
    pull(relative_call_count)
generate_macro("PercCallerEnvSubstituteCalls", PercCallerEnvSubstituteCalls)

PercNewEnvSubstituteCalls <-
    substitute_class_summary %>%
    filter(substitute_class == "NewScope") %>%
    pull(relative_call_count)
generate_macro("PercNewEnvSubstituteCalls", PercNewEnvSubstituteCalls)

CountDynamicEnvSubstituteCalls <-
    substitute_class_summary %>%
    filter(substitute_class %in% c("DynamicScope", "StaticScope")) %>%
    pull(call_count) %>%
    sum()
generate_macro("CountDynamicEnvSubstituteCalls", CountDynamicEnvSubstituteCalls)

@@

```{r, eval = TRUE, echo = FALSE}
alphabetical_row_names = c(letters, toupper(letters))

datatable(promise_lifecycle_summary)

CountLifecycleTotal <-
    promise_lifecycle_summary %>% 
    pull(total_lifecycle_count) %>%
    first()

generate_macro("CountLifecycleTotal", CountLifecycleTotal)

CountLifecycleUnescapedArguments <-
    promise_lifecycle_summary %>% 
    filter(argument, !escaped) %>% 
    pull(lifecycle_count)

generate_macro("CountLifecycleUnescapedArguments", CountLifecycleUnescapedArguments)

CountLifecycleEscapedArguments <-
    promise_lifecycle_summary %>% 
    filter(argument, escaped) %>% 
    pull(lifecycle_count)

generate_macro("CountLifecycleEscapedArguments", CountLifecycleEscapedArguments)

CountLifecycleNonArguments <-
    promise_lifecycle_summary %>% 
    filter(!argument) %>% 
    pull(lifecycle_count)

generate_macro("CountLifecycleNonArguments", CountLifecycleNonArguments)

decode_sequence <- function(seq) {
        unlist(str_split(str_sub(seq, 2, -2), " "))
}

expand_lifecycle <- function(action, count) {
    if(action == "()") {
        ""
    } else {
        action_seq <- decode_sequence(action)
        count_seq <- decode_sequence(count)
        str_c(rep(action_seq, count_seq), collapse = " ")
    }
}

promise_lifecycle_list <-
    promise_lifecycle_list %>%
    mutate(event = map2_chr(action, count, expand_lifecycle)) %>%
    select(-action, -count, -promise_count)

escaped_argument_list <-
    promise_lifecycle_list %>%
    filter(argument, escaped)

datatable(escaped_argument_list)

escaped_argument_list %>%
    select(event, relative_promise_count) %>%
    arrange(desc(relative_promise_count)) %>%
    slice(1:10) %>%
    mutate(index = alphabetical_row_names[dplyr::row_number()]) %>%
    pmap(function(event, relative_promise_count, index) {
        name <- str_c("TextLifecycleEscapedArguments", index)
        generate_macro(name, event)
        name <- str_c("PercLifecycleEscapedArguments", index)
        generate_macro(name, relative_promise_count)
    })

unescaped_argument_list <-
    promise_lifecycle_list %>%
    filter(argument, !escaped)

datatable(unescaped_argument_list)

unescaped_argument_list %>%
    select(event, relative_promise_count) %>%
    arrange(desc(relative_promise_count)) %>%
    slice(1:10) %>%
    mutate(index = alphabetical_row_names[dplyr::row_number()]) %>%
    pmap(function(event, relative_promise_count, index) {
        name <- str_c("TextLifecycleUnescapedArguments", index)
        generate_macro(name, event)
        name <- str_c("PercLifecycleUnescapedArguments", index)
        generate_macro(name, relative_promise_count)
    })

non_argument_list <-
    promise_lifecycle_list %>%
    filter(!argument)

datatable(non_argument_list)

non_argument_list %>%
    select(event, relative_promise_count) %>%
    arrange(desc(relative_promise_count)) %>%
    slice(1:10) %>%
    mutate(index = alphabetical_row_names[dplyr::row_number()]) %>%
    pmap(function(event, relative_promise_count, index) {
        name <- str_c("TextLifecycleNonArguments", index)
        generate_macro(name, event)
        name <- str_c("PercLifecycleNonArguments", index)
        generate_macro(name, relative_promise_count)
    })

@@

context_sensitive_binding_lookup <- 
    context_sensitive_lookup_summary %>%
    pull(binding_lookup) %>%
    sum()
generate_macro("CountConSenBndgLkp", context_sensitive_binding_lookup)

context_sensitive_argument_binding_lookup_failure <- 
    context_sensitive_lookup_summary %>%
    filter(argument & !success) %>%
    pull(binding_lookup) %>%
    sum()
generate_macro("CountConSenArgBndgLkpFailure", context_sensitive_argument_binding_lookup_failure)

context_sensitive_argument_force <-
    context_sensitive_lookup_summary %>% 
    filter(argument) %>% 
    pull(forced) %>% 
    sum()
generate_macro("CountConSenArgForce", context_sensitive_argument_force)

context_sensitive_argument_force_successful <-
    context_sensitive_lookup_summary %>% 
    filter(argument & success) %>% 
    pull(forced) %>% 
    sum()
generate_macro("CountConSenArgForceSuccess", context_sensitive_argument_force_successful)
generate_macro("PercConSenArgForceSuccess", context_sensitive_argument_force_successful/context_sensitive_argument_force)

context_sensitive_argument_force_failure <-
    context_sensitive_lookup_summary %>% 
    filter(argument & !success) %>% 
    pull(forced) %>% 
    sum()
generate_macro("CountConSenArgForceFailure", context_sensitive_argument_force_failure)
generate_macro("PercConSenArgForceFailure", context_sensitive_argument_force_failure / context_sensitive_argument_force)

context_sensitive_lookup_symbols <-
    context_sensitive_lookup_symbols %>%
    filter(formal_parameter_position != -1)

symbol_overload_table <-
    context_sensitive_lookup_symbols %>%
    filter(argument) %>%
    group_by(symbol) %>%
    summarize(count = n_distinct(function_id, formal_parameter_position)) %>%
    ungroup() %>%
    arrange(desc(count))

context_sensitive_overloaded_parameter_count <-
    context_sensitive_lookup_symbols %>%
    distinct(function_id, formal_parameter_position) %>%
    nrow()
generate_macro("CountConSenParameter", context_sensitive_overloaded_parameter_count)

context_sensitive_overloaded_function_count <-
    context_sensitive_lookup_symbols %>%
    distinct(function_id) %>%
    nrow()
generate_macro("CountConSenFunction", context_sensitive_overloaded_function_count)

row_names <- c(letters, toupper(letters))

symbol_overload_table %>%
    slice(1:30) %>%
    mutate(index = row_names[dplyr::row_number()]) %>%
    pmap(function(index, symbol, count) {
        generate_macro(paste0("CodeConSenSymbol", index), symbol)
        generate_macro(paste0("CountConSenSymbol", index), count)
    })

datatable(symbol_overload_table)
```

@@

AllGcPromises <-
    promise_gc_summary %>%
    pull(promise_count) %>%
    sum()

CountSingleGcPromises <-
    promise_gc_summary %>%
    filter(single_gc_cycle) %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercSingleGCPromises", CountSingleGcPromises/AllGcPromises)

CountMultipleGcPromises <-
    promise_gc_summary %>%
    filter(!single_gc_cycle) %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercMultipleGcPromises", CountMultipleGcPromises/AllGcPromises)

CountMultipleGcEscapedArgumentPromises <-
    promise_gc_summary %>%
    filter(!single_gc_cycle & escaped & argument) %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercMultipleGcEscapedArgumentPromises", CountMultipleGcEscapedArgumentPromises/CountMultipleGcPromises)

CountMultipleGcUnescapedArgumentPromises <-
    promise_gc_summary %>%
    filter(!single_gc_cycle & argument & !escaped) %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercMultipleGcUnescapedArgumentPromises", CountMultipleGcUnescapedArgumentPromises/CountMultipleGcPromises)

CountMultipleGcNonArgumentPromises <-
    promise_gc_summary %>%
    filter(!single_gc_cycle & !argument) %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercMultipleGcNonArgumentPromises", CountMultipleGcNonArgumentPromises/CountMultipleGcPromises)

@@

```{r, eval = TRUE, echo = FALSE}

CountEffectPromises <-
    created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("CountEffectPromises", CountEffectPromises)

CountEffectSamePromises <-
    created_side_effects %>%
    filter(mode == "S") %>%
    pull(count) %>%
    sum()

generate_macro("CountEffectSamePromises", CountEffectSamePromises)
generate_macro("PercEffectSamePromises", CountEffectSamePromises / CountEffectPromises)

CountEffectOtherPromises <-
    created_side_effects %>%
    filter(mode == "O") %>%
    pull(count) %>%
    sum()

generate_macro("CountEffectOtherPromises", CountEffectOtherPromises)
generate_macro("PercEffectOtherPromises", CountEffectOtherPromises / CountEffectPromises)

CountEffectLexicalPromises <-
    created_side_effects %>%
    filter(mode == "L") %>%
    pull(count) %>%
    sum()

generate_macro("CountEffectLexicalPromises", CountEffectLexicalPromises)
generate_macro("PercEffectLexicalPromises", CountEffectLexicalPromises / CountEffectPromises)
```
## Direct & Same Environment

```{r, eval = TRUE, echo = FALSE}

direct_S_created_side_effects <-
    created_side_effects %>%
    filter(direct & mode == "S") %>%
    arrange(desc(count))

datatable(direct_S_created_side_effects)

CountEffectDirectSamePromises <-
    direct_S_created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("PercEffectDirectSamePromises", CountEffectDirectSamePromises / CountEffectSamePromises)
```

## Indirect & Same Environment

```{r, eval = TRUE, echo = FALSE}
indirect_S_created_side_effects <-
    created_side_effects %>%
    filter(!direct & mode == "S") %>%
    arrange(desc(count))

datatable(indirect_S_created_side_effects)

CountEffectIndirectSamePromises <-
    indirect_S_created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("PercEffectIndirectSamePromises", CountEffectIndirectSamePromises / CountEffectSamePromises)
```

## Direct & Other Environment

```{r, eval = TRUE, echo = FALSE}
direct_O_created_side_effects <-
    created_side_effects %>%
    filter(direct & mode == "O") %>%
    arrange(desc(count))

datatable(direct_O_created_side_effects)

CountEffectDirectOtherPromises <-
    direct_O_created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("PercEffectDirectOtherPromises", CountEffectDirectOtherPromises / CountEffectOtherPromises)
```

## Indirect & Other Environment

```{r, eval = TRUE, echo = FALSE}
indirect_O_created_side_effects <-
    created_side_effects %>%
    filter(!direct & mode == "O") %>%
    arrange(desc(count))

datatable(indirect_O_created_side_effects)

CountEffectIndirectOtherPromises <-
    indirect_O_created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("PercEffectIndirectOtherPromises", CountEffectIndirectOtherPromises / CountEffectOtherPromises)
```

## Direct & Lexical Parent Environment

```{r, eval = TRUE, echo = FALSE}
direct_L_created_side_effects <-
    created_side_effects %>%
    filter(direct & mode == "L") %>%
    arrange(desc(count))

datatable(direct_L_created_side_effects)

CountEffectDirectLexicalPromises <-
    direct_L_created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("PercEffectDirectLexicalPromises", CountEffectDirectLexicalPromises / CountEffectLexicalPromises)
```

## Indirect & Lexical Parent Environment

```{r, eval = TRUE, echo = FALSE}
indirect_L_created_side_effects <-
    created_side_effects %>%
    filter(!direct & mode == "L") %>%
    arrange(desc(count))

datatable(indirect_L_created_side_effects)

CountEffectIndirectLexicalPromises <-
    indirect_L_created_side_effects %>%
    pull(count) %>%
    sum()

generate_macro("PercEffectIndirectLexicalPromises", CountEffectIndirectLexicalPromises / CountEffectLexicalPromises)
```
##datatable(observed_side_effects)
```

@@

```{r, eval = TRUE, echo = FALSE}

datatable(argument_promise_count_by_call_depth)

## remove non-argument and escaped promises
argument_promise_count_by_call_depth <-
    argument_promise_count_by_call_depth %>%
    filter(call_depth != "-1")

CountPromiseAllDepth <-
    argument_promise_count_by_call_depth %>%
    pull(promise_count) %>%
    sum()

CountPromiseZeroDepth <-
    argument_promise_count_by_call_depth %>%
    filter(call_depth == "0") %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercPromiseZeroDepth", CountPromiseZeroDepth / CountPromiseAllDepth)

CountPromiseOneDepth <-
    argument_promise_count_by_call_depth %>%
    filter(call_depth == "1") %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercPromiseOneDepth", CountPromiseOneDepth / CountPromiseAllDepth)

CountPromiseMultipleDepth <-
    argument_promise_count_by_call_depth %>%
    filter(call_depth != "0" & call_depth != "1") %>%
    pull(promise_count) %>%
    sum()
generate_macro("PercPromiseMultipleDepth", CountPromiseMultipleDepth / CountPromiseAllDepth)

@@

save_local_graph(argument_promise_call_depth_graph, "forcedepth.pdf", width=12, height=2.5)
```

@@

```{r, eval = TRUE, echo = FALSE}
datatable(promise_argument_forced_by_promise_argument)

CountMutualForcing <-
    sum(promise_argument_forced_by_promise_argument$count)

generate_macro("CountMutualForcing", CountMutualForcing)

CountMutualForcingFunctions <-
    n_distinct(promise_argument_forced_by_promise_argument$function_id)

generate_macro("CountMutualForcingFunctions", CountMutualForcingFunctions)

CountMutualForcingArgumentPositions <-
    n_distinct(promise_argument_forced_by_promise_argument$function_id, 
               promise_argument_forced_by_promise_argument$forcing_actual_argument_position)

generate_macro("CountMutualForcingArgumentPositions", CountMutualForcingArgumentPositions)

CountMutualForcedArgumentPositions <-
    n_distinct(promise_argument_forced_by_promise_argument$function_id, 
               promise_argument_forced_by_promise_argument$formal_parameter_position)

generate_macro("CountMutualForcedArgumentPositions", CountMutualForcedArgumentPositions)
```

@@

```{r, eval = TRUE, echo = FALSE}

s3_promise_force <-
    argument_promise_count_by_forcing_scope %>%
    filter(forcing_scope == "S3") %>%
    pull(promise_count) %>%
    sum()
generate_macro("CountSThreePromiseForce", s3_promise_force)

s4_promise_force <-
    argument_promise_count_by_forcing_scope %>%
    filter(forcing_scope == "S4") %>%
    pull(promise_count) %>%
    sum()
generate_macro("CountSFourPromiseForce", s4_promise_force)

datatable(argument_promise_count_by_forcing_scope)
```

### 2019-08-15 report/paper.Rmd
Update reports and add new analysis

PercDispatch <- 
    argument_promise_count_by_dispatch_type %>% 
    filter(dispatch_type != "None") %>% 
    pull(relative_promise_count) %>% 
    sum()
generate_macro("PercDispatch", PercDispatch)
```

@@

@@

```{r eval = TRUE, echo = FALSE}
CountEscArg <- sum(escaped_argument_count_by_nature$argument_count)
```

Of the `r format_as_absolute(CountPromiseArgument)` promises we have
observed, only `r format_as_absolute(CountEscArg)` escape. This is a rather small number, and we
need to establish the reason why the promise escape. One thing we can
measure is the return type of the functions for which promises escape.

>>>>

PercDispatch <- 
    argument_promise_count_by_dispatch_type %>% 
    filter(dispatch_type != "None") %>% 
    pull(relative_promise_count) %>% 
    sum()
generate_macro("PercDispatch", PercDispatch)

```

@@

CountArgumentPromiseExpressionFunctionCall <-
    argument_promise_count_by_expression_type %>%
    filter(type == "Function Call") %>%
    pull(count) %>%
    sum()

generate_macro("PercArgumentPromiseExpressionFunctionCall", CountArgumentPromiseExpressionFunctionCall/total_count)

@@

```{r eval = TRUE, echo = FALSE}
CountEscapeArgument <- sum(escaped_argument_count_by_nature$argument_count)
generate_macro("CountEscapeArgument", CountEscapeArgument)
```

Of the `r format_as_absolute(CountPromiseArgument)` promises we have
observed, only `r format_as_absolute(CountEscapeArgument)` escape. This is a rather small number, and we
need to establish the reason why the promise escape. One thing we can
measure is the return type of the functions for which promises escape.
